---
title: "Reproducible Research: Peer Assessment 1"
author: "WendyD"
output: 
  html_document:
    keep_md: true
---

## Loading and preprocessing the data
first load the data
```{r, echo=TRUE}
data1<-read.csv("activity.csv", header = TRUE, sep = ",")
str(data1)
```

see how many rows the data has
```{r, echo=TRUE}
nrow(data1)
```

## What is mean total number of steps taken per day?
```{r, echo=TRUE}
sumstep <- aggregate(data1$steps~data1$date,data= data1, FUN = sum)
colnames(sumstep)<- c("date", "steps")
```


Making a histogram for the total number of steps taken per day
```{r, echo=TRUE}
hist(sumstep$steps, main = "Total Steps taken each day", xlab="Number Steps")
```

The mean total number of steps taken per day
```{r, echo=TRUE}
meanstep<- mean(sumstep$steps)
meanstep
medianstep<-median(sumstep$steps)
medianstep

```


## What is the average daily activity pattern?
Make a time series plot of the 5-minute interval(x axis) and the average number of steps taken, averaged across all days (y-axis)

```{r, echo=TRUE}
meanstep_interval <- aggregate(data1$steps~data1$interval, data=data1, mean)
```

Assigning name to the new data frame containing average steps by time interval
```{r, echo=TRUE}
colnames(meanstep_interval)<-c("interval", "steps")
```

Plot graph showing average steps across different interval (X axis)
```{r, echo=TRUE}
plot(meanstep_interval$interval, meanstep_interval$steps, type="l",xlab="Interval", ylab="Steps", main= "Average Daily Activity Pattern ")
```
Whoch 5 minute time interval contains the most step
```{r, echo=TRUE}
max_step<-max(meanstep_interval$steps)
max_interval<-meanstep_interval$interval[meanstep_interval$steps == max_step]
max_interval
```
The 5 minute interval where there is max of average steps taken is `r max_interval`

## Imputing missing values
Calculate the number of row with NA values
```{r, echo=TRUE}
totalNA<-sum(is.na(data1$steps))
totalNA
```
Replacing NA value with mean value for that same interval (meanstep_interval)
first make a new data so we can retrace our mistakes if needed
```{r, echo=TRUE}
data2<-data1
```
Now replacing the NA value with mean steps for each time interval
```{r, echo=TRUE}
NAlocation<-which(is.na(data2$steps))
for (i in NAlocation) {
    data2$steps[i]<-with(meanstep_interval, steps[interval ==data2$interval[i]])
}

```
Now test if we replace all NA or not
```{r, echo=TRUE}
sum(is.na(data2))
```
Making histogram of total steps taken per day. Report mean and median
First we calculate the total of steps taken by day
```{r, echo=TRUE}
sumstep2<-aggregate(data2$steps~data2$date, data=data2, sum)
colnames(sumstep2)<- c("date", "steps")

```
Plug in a histogram

```{r, echo=TRUE}
hist(sumstep2$steps,xlab = "Total Steps", main = "Total steps taken each day")
```

The mean and mediantotal number of steps taken per day

```{r, echo=TRUE}
meanstep2<- mean(sumstep2$steps)
meanstep2
medianstep2<-median(sumstep2$steps)
medianstep2
```

To compared with the mean and median values from the old data
We make a new data frame
```{r, echo=TRUE}
compare<-data.frame(meancol=c(meanstep, meanstep2), mediancol=c(medianstep, medianstep2))
compare
```
From the table, we see that after we fill all the NA value with the average step takens per interval, we have the same mean and median values



## Are there differences in activity patterns between weekdays and weekends?


